## General Social Survey - highest_year_of_school_completed

```{r}
library(tidyverse)
library(Stat2Data)
library(skimr)
library(leaps)
library(ggplot2)
```

```{r}
GSS <-read_csv("GSS_clean.csv")
```

**GSS** stands for General Social Survey.

```{r}
names(GSS)
```

**Response Variable:**

highest_year_of_school_completed

**Explanatory Variables:**

1.  highest_year_school_completed_father
2.  highest_year_school_completed_mother
3.  born_in_us
4.  age_of_respondent
5.  number_of_brothers_and sisters
6.  respondents_sex
7.  marital_status

Separating the variables of concern in another dataset:

```{r}
newGSS <- select(GSS, respondents_sex, highest_year_school_completed_mother, age_of_respondent, marital_status, number_of_brothers_and_sisters, born_in_us, highest_year_school_completed_father, highest_year_of_school_completed)
```

```{r}
skim(newGSS)
summary(newGSS)
```

There seem to be plenty of NA values. Since I have 2348 observations, I can afford to remove all NA values.

```{r}
newGSS<- na.omit(newGSS)
```

I will conduct a backward elimination method to determine which variables I will keep and which ones I will remove. Since there are only 7 variables, I will conduct this manually.

```{r}
 model<- lm(highest_year_of_school_completed ~ ., data = newGSS)
 summary(model)
```

It seems like the model is taking each age under the age_of_respondent variable as a separate category.

```{r}
summary(newGSS$age_of_respondent)

```

It seems like the entries for age_of_respondent variable are String or character. Therefore, I am converting them to numeric.

```{r}
newGSS <- newGSS %>%
  mutate(age_numeric = parse_number(age_of_respondent))
```

To check:

```{r}
summary(newGSS$age_numeric)
```

It has now been converted to numeric.

```{r}
model1<- lm(highest_year_of_school_completed ~ respondents_sex+ highest_year_school_completed_mother+ number_of_brothers_and_sisters + marital_status + born_in_us + highest_year_school_completed_father + age_numeric,  data = newGSS)
 summary(model1)
```

It seems like marital_status is not a good predictor of highest_year_of_school_completed, thus, I am eliminating that.

```{r}
model2<- lm(highest_year_of_school_completed ~ respondents_sex+ highest_year_school_completed_mother+ number_of_brothers_and_sisters +  born_in_us + highest_year_school_completed_father + age_numeric,  data = newGSS)
 summary(model2)
```

It seems like born_in_us is not a good predictor, and neither is respondents_sex, however, born_in_us has a larger p-value, thus I will eliminate that first.

```{r}
model3<- lm(highest_year_of_school_completed ~ respondents_sex+ highest_year_school_completed_mother+ number_of_brothers_and_sisters  + highest_year_school_completed_father + age_numeric,  data = newGSS)
 summary(model3)
```

It seems like respondents_sex is the only non-significant predictor that needs to be eliminated. Since all these eliminations does not impact the adjusted R\^2 greatly, I don't think there is any problem in eliminating them.

```{r}
model3<- lm(highest_year_of_school_completed ~ highest_year_school_completed_mother+ number_of_brothers_and_sisters  + highest_year_school_completed_father + age_numeric,  data = newGSS)
 summary(model3)
```

It seems like all the variables now are significant predictors of highest_year_of_school_completed.

Checking the overall age distribution in the dataset.

```{r}
summary(newGSS$age_numeric)
summary(newGSS$age_numeric < 25)
```

The age of the respondents ranges between 18 to 89 and there are 130 individuals with age under 25.

Equation:

ŷ = 9.25334 + 0.15343(highest_year_school_completed_mother) - 0.12340 (number_of_brothers_and_sisters ) + 0.20069 (highest_year_school_completed_father) + 0.02315(age_numeric).

where ŷ is highest_year_of_school_completed-hat.

Checking whether or not the conditions for inference are being met.

1.  Equality of Variance and Linearity:

```{r}
plot(model3, which = 1)
```

2.  Normality:

    ```{r}
    plot(model3, which = 2)
    ```

#### 
